\documentclass[main.tex]{subfiles}

\begin{document}

\section{2D Motion Application}

Two-dimensional motion breaks down into two parts: projectile motion and circular motion.

\subsection{Projectile Motion} \label{tMF7JY}

\Gls{projectile motion} is the motion of an object thrown (projected) into the air when, after the initial force that launches the object, air resistance is negligible and the only other force that object experiences is the force of gravity. The object is called a \gls{projectile}, and its path is called its \gls{trajectory}. \Gls{air resistance} is a frictional force that slows its motion and can significantly alter the trajectory of the motion. Due to the difficulty in calculation, only situations in which the deviation from projectile motion is negligible and air resistance can be ignored are considered in introductory physics. That approximation is often quite accurate.

\vspace{1em}

The most important concept in projectile motion is that when air resistance is ignored, horizontal and vertical motions are independent, meaning that they don't influence one another. Figure 5.27 compares a cannonball in free fall (in blue) to a cannonball launched horizontally in projectile motion (in red). You can see that the cannonball in free fall falls at the same rate as the cannonball in projectile motion. Keep in mind that if the cannon launched the ball with any vertical component to the velocity, the vertical displacements would not line up perfectly.

\vspace{1em}

Since vertical and horizontal motions are independent, we can analyze them separately, along perpendicular axes. To do this, we separate projectile motion into the two components of its motion, one along the horizontal axis and the other along the vertical.

\begin{center}
    \begin{tikzpicture} 
    \tikzmath{
        \vi = 28.6;
        \yi = 60;
        \thetai = 0.0;
    }
    \tikzset{declare function={f(\x)=\yi + tan(\thetai)*\x - 9.8*\x^2/(2*(\vi*cos(\thetai))^2);}} %equation of path
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=8cm,height=8cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax=120,
        ymin=0, ymax=80,
        ]
        \draw[dashed,domain=0:100,variable=\x,samples=200] plot ({\x},{f(\x)});
        \draw[dashed] (0,{f(0)}) -- ++(100,0);
        \draw[dashed] (0,{f(0)}) -- ++(0,-{f(0)});
        
        \fill (0,{f(0)}) circle (2pt);
        \fill (20,{f(20)}) circle (2pt); 
        \fill (40,{f(40)}) circle (2pt); 
        \fill (60,{f(60)}) circle (2pt); 
        \fill (80,{f(80)}) circle (2pt); 
        \fill (100,{f(100)}) circle (2pt); 
        
        \draw[dashed] (20,{f(0)}) -- ++(0,{-(f(0)-f(20))});
        \draw[dashed] (40,{f(0)}) -- ++(0,{-(f(0)-f(40))});
        \draw[dashed] (60,{f(0)}) -- ++(0,{-(f(0)-f(60))});
        \draw[dashed] (80,{f(0)}) -- ++(0,{-(f(0)-f(80))});
        \draw[dashed] (100,{f(0)}) -- ++(0,{-(f(0)-f(100))});
        \fill[cyan] (20,{f(0)}) circle (2pt);
        \fill[cyan] (40,{f(0)}) circle (2pt);
        \fill[cyan] (60,{f(0)}) circle (2pt);
        \fill[cyan] (80,{f(0)}) circle (2pt);
        \fill[cyan] (100,{f(0)}) circle (2pt);
        
        \draw[dashed] (0,{f(20)}) -- ++(20,0);
        \draw[dashed] (0,{f(40)}) -- ++(40,0);
        \draw[dashed] (0,{f(60)}) -- ++(60,0);
        \draw[dashed] (0,{f(80)}) -- ++(80,0);
        \draw[dashed] (0,{f(100)}) -- ++(100,0);
        \fill[red] (0,{f(20)}) circle (2pt);
        \fill[red] (0,{f(40)}) circle (2pt);
        \fill[red] (0,{f(60)}) circle (2pt);
        \fill[red] (0,{f(80)}) circle (2pt);
        \fill[red] (0,{f(100)}) circle (2pt);
        \end{axis}
    \end{tikzpicture}
    \captionsetup{type=figure,margin=1in,font=scriptsize}
    \captionof{figure}{The diagram shows the projectile motion of a cannonball shot at a horizontal angle versus one dropped with no horizontal velocity. Note that both cannonballs have the same vertical position over time.}
\end{center}

\vspace{1em}

We'll call the horizontal axis the $x$-axis and the vertical axis the $y$-axis. For notation, $d$ is the total displacement, and $x$ and $y$ are its components along the horizontal and vertical axes. The magnitudes of these vectors are $x$ and $y$, as illustrated in Figure \ref{i1Y7vV}.

\begin{center}
    \begin{tikzpicture} 
    \tikzmath{
        \vi = 18.0;
        \yi = 0;
        \thetai = 70.0;
    }
    \tikzset{declare function={f(\x)=\yi + tan(\thetai)*\x - 9.8*\x^2/(2*(\vi*cos(\thetai))^2);}} %equation of path
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=8cm,height=8cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax=20,
        ymin=0, ymax=18,
        ]
        \node at (15.6,0.35) {$\square$};
        \draw[dashed,domain=0:16,variable=\x,samples=200] plot ({\x},{f(\x)}) node[above left=-5pt] {\faSoccerBallO};
        \draw[->,thick] (0,0) -- ++(16,{f(16)}) node[pos=0.5,above=3pt]{$d$};
        \draw[->,dashed,thick,cyan] (0,0) -- ++(16,0) node[pos=0.5,below] {$x$};
        \draw[->,dashed,thick,red] (16,0) -- ++(0,{f(16)}) node[pos=0.5,right] {$y$};
        \node at (2.5,0.65) {$\theta$};
        \draw (1.5,0) arc (0:35:1.5);
        \end{axis}
    \end{tikzpicture}
    \captionsetup{type=figure,margin=1in,font=scriptsize}
    \captionof{figure}{A boy kicks a ball at angle $\theta$, and it is displaced a distance of $d$ along its trajectory.}
    \label{i1Y7vV}
\end{center}

\vspace{1em}

As usual, we use velocity, acceleration, and displacement to describe motion. We must also find the components of these variables along the $x$- and $y$-axes. The components of acceleration are then very simple: $a_y = -g = -\SI{9.80}{m/s^2}$. Note that this definition defines the upwards direction as positive. Because gravity is vertical, $a_x = 0$. Both accelerations are constant, so we can use the kinematic equations. For review, the kinematic equations from a previous chapter are summarized below.

\begin{align}
    x &= x_0 + v_{\text{avg}} t\\[0.5ex]
    v_{\text{avg}} &= \frac{v_0 + v}{2}\\[0.5ex]
    v &= v_0 + a t\\[0.5ex]
    x &= x_0 + v_0 t + \frac{1}{2} a t^2\\[0.5ex]
    v^2 &= v_0^2 + 2 a (x - x_0)
\end{align}

where $x$ is position, $x_0$ is initial position, $v$ is velocity, $v_avg$ is average velocity, $t$ is time, and $a$ is acceleration.

\vspace{1em}

The following four steps are used to analyze projectile motion:

\vspace{1em}

\textbf{Step 1}: Separate the motion into horizontal and vertical components along the $x$- and $y$-axes. These axes are perpendicular, so the Pythagorean theorem, $A^2 = A_x^2 + A_y^2$ is used. The magnitudes of the displacement $d$ along $x$- and $y$-axes are called $x$ and  $y$. The magnitude of the velocity is $v$, and the magnitudes of the components of the velocity, $v_x$ and $v_y$, relate to $v$ also by the Pythagorean theorem. Initial values are denoted with a subscript 0.

\vspace{1em}

\textbf{Step 2}: Treat the motion as two independent one dimensional motions, one horizontal and the other vertical. The kinematic equations for horizontal motion take the following forms under the condition that horizontal acceleration is zero ($a_x = 0$):

\begin{align}
    x &= x_0 + v_x t \label{7fMAPG}\\[0.5ex]
    v_x &= v_{0x} = \text{velocity is a constant} \label{dCvDf4}
\end{align}

For vertical motion, the kinematic equations take the following forms under the condition that vertical acceleration is $a_y = -g = \SI{-9.80}{m/s^2}$:

\begin{align}
    y &= y_0 + \frac{1}{2}\left(v_{0y} + v_y\right)t\\[0.5ex]
    v_y &= v_{0y} - gt \label{SjYaoE} \\[0.5ex]
    y &= y_0 + v_{0y}t - \frac{1}{2}  g t^2  \label{36YuvF} \\[0.5ex]
    v_y^2 &= v_{0y}^2 - 2 g (y - y_0) \label{wROSXN}
\end{align}

\vspace{1em}

\textbf{Step 3}: Solve for the unknowns in the two separate motions (one horizontal and one vertical). Note that the only common variable between the motions is time $t$. The problem solving procedures here are the same as for one-dimensional kinematics.

\vspace{1em}

\textbf{Step 4}: Recombine the two motions to find the total displacement $d$ and velocity $v$. These are related to their components via the Pythagorean theorem as shown below:

\begin{align}
    d &= \sqrt{x^2 + y^2}\\[0.5ex]
    v &= \sqrt{v_x^2 + v_y^2}
\end{align}
  .

\begin{center}
    \begin{tikzpicture}[
        declare function={f(\x)=\yi + tan(\thetai)*\x - 10*\x^2/(2*(\vi*cos(\thetai))^2);},
        declare function={t(\x)=\x/(\vi*cos(\thetai));}
    ]
    \tikzmath{
        \grav = 10;
        \vi = 20;
        \yi = 0;
        \thetai = 45.0;
        \viy = \vi*sin(\thetai);
        \ymax = (\viy^2)/(2*\grav);
        \tmax = 2*\ymax/\viy;
        \vix = \vi*cos(\thetai);
        \xmax = \vix*\tmax;
        \sf = 0.3; %scale factor for vector components
    }
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=12cm,height=12cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax=50,
        ymin=0, ymax=15,
        ]
        
        \draw[fill=black] (0,0) circle (2pt);
        \draw[thick,violet,->] (0,0) -- ++(\vix*\sf,\viy*\sf) node[below right] {$v_0$};
        \draw[thick,dashed,red,->] (0,0) -- ++(0,{\sf*\vi*sin(\thetai)}) node[right=-2pt] {$v_{0y}$};
        \draw[thick,dashed,cyan,->] (0,0) -- ++(\vix*\sf,0) node[pos=1.3, above=-2pt] {$v_{0x}$};
        \draw (1,0) arc (0:75:0.6) node[midway,right=-1pt] {$\theta_0$};
        
        \pgfplotsinvokeforeach{5,10,15}{
                \draw[fill=black] (#1,{f(#1)}) circle (2pt);
                \draw[thick,dashed,red,->] (#1,{f(#1)}) -- ++(0,{\sf*(\vi*sin(\thetai) - \grav*t(#1))}) node[above=-2pt] {$v_y$};
                \draw[thick,dashed,cyan,->] (#1,{f(#1)}) -- ++(\vix*\sf,0) node[right=-2pt] {$v_x$};
                \draw[thick,violet,->] (#1,{f(#1)}) -- ++(\vix*\sf,{\sf*(\vi*sin(\thetai) - \grav*t(#1))}) node[above=-2pt] {$v$};
                }
                
        \draw[fill=black] (\xmax,\ymax) circle (2pt);
        \draw[thick,violet,->] (\xmax,\ymax) -- ++(\vix*\sf,0) node[right=-2pt] {$v = v_x$};
        
        \pgfplotsinvokeforeach{25,30,35,45}{
                \draw[fill=black] (#1,{f(#1)}) circle (2pt);
                \draw[thick,dashed,red,->] (#1,{f(#1)}) -- ++(0,{\sf*(\vi*sin(\thetai) - \grav*t(#1))}) node[below=-2pt] {$v_y$};
                \draw[thick,dashed,cyan,->] (#1,{f(#1)}) -- ++(\vix*\sf,0) node[right=-2pt] {$v_x$};
                \draw[thick,violet,->] (#1,{f(#1)}) -- ++(\vix*\sf,{\sf*(\vi*sin(\thetai) - \grav*t(#1))}) node[below=-2pt] {$v$};
                }

                \draw[fill=black] (40,{f(40)}) circle (2pt);
                \draw[thick,dashed,red,->] (40,{f(40)}) -- ++(0,{\sf*(\vi*sin(\thetai) - \grav*t(40))}) node[left=-2pt] {$v_y=-v_{0y}$};
                \draw[thick,dashed,cyan,->] (40,{f(40)}) -- ++(\vix*\sf,0) node[above=-2pt] {$v_x$};
                \draw[thick,violet,->] (40,{f(40)}) -- ++(\vix*\sf,{\sf*(\vi*sin(\thetai) - \grav*t(40))}) node[below=-2pt] {$v$};
                \draw (40.8,0) arc (0:-\thetai:0.8) node[midway,right=-1pt] {$\theta =-\theta_0$};

                \draw (5.8,{f(5)}) arc (0:40:0.8) node[midway,right] {$\theta$};
        \end{axis}
    \end{tikzpicture}
    \captionsetup{type=figure,font=scriptsize}
    \captionof{figure}{We analyze two-dimensional projectile motion by breaking it into two independent one-dimensional motions along the vertical and horizontal axes. The horizontal motion is constant. The velocity in the vertical direction begins to decrease, then becomes zero at the peak, then increases in the downward direction.}
    \label{xRbMxN}
\end{center}

\clearpage

\begin{example} \label{4Dn9PW}
    During a fireworks display like the one illustrated in Figure \ref{ZxjpEr}, a shell is shot into the air with an initial speed of \SI{70.0}{m/s} at an angle of \ang{75} above the horizontal. The fuse is timed to ignite the shell just as it reaches its highest point above the ground. Calculate the height at which the shell explodes. (This peak vertical position is defined as \gls{maximum height}.)
\end{example}

\begin{center}
    \begin{tikzpicture}[
        declare function={f(\x)=\yi + tan(\thetai)*\x - \grav*\x^2/(2*(\vi*cos(\thetai))^2);},
        declare function={t(\x)=\x/(\vi*cos(\thetai));}
    ]
    \tikzmath{
        \grav = 9.8;
        \vi = 70;
        \yi = 0.0;
        \thetai = 75.0;
        \viy = \vi*sin(\thetai);
        \ymax = (\viy^2)/(2*\grav);
        \tmax = 2*\ymax/\viy;
        \vix = \vi*cos(\thetai);
        \xmax = \vix*\tmax;
        \sf = 0.25; %scale factor for vector components
    }
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=8cm,height=8cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax=175,
        ymin=0, ymax=275,
        ]
        \draw[domain=0:\xmax,variable=\x,samples=250] plot ({\x},{f(\x)});
        \draw (10,0) arc (0:35:30) node[pos=0.8,right=2pt] {$\theta = \ang{75}$};
        \draw[dashed] (0,\ymax) node[left] {$h$} -- ++(\xmax,0);
        \draw[dashed] (\xmax,0) -- ++(0,\ymax);
        \draw[ultra thick,violet,->] (0,0) -- ++({\vi*cos(\thetai)},{\vi*sin(\thetai)}) node[left] {$v$};
        \end{axis}
    \end{tikzpicture}
    \captionsetup{type=figure,margin=1in,font=scriptsize}
    \captionof{figure}{The diagram shows the trajectory of a fireworks shell.}
    \label{ZxjpEr}
\end{center}

\Solution We are given the initial speed and launch angle: $v_0 = \SI{70}{m/s}$ and $\theta_0 = \ang{75}$. We know the magnitude of the acceleration due to gravity is $g = \SI{9.8}{m/s^2}$. Since horizontal and vertical motions are independent, we start by analyzing only the vertical motion. The $y$-component of the initial velocity is given by

\begin{equation*}
    v_{0y} = v_0 \sin{\theta_0} = 70 \sin{(\ang{75})} = 67.61
\end{equation*}

If we assume the firework starts at the origin, its initial vertical position is zero: $y_0 = 0$.

\vspace{1em}

The unknown we are looking for is final vertical position (or height above $y_0$): $y =\ ?$ 

\vspace{1em}

The key to solving this problem is the following fact: From Figure ?.?? we see that at the highest point in the trajectory, the $y$-component of velocity is zero: $v_y = 0$. Therefore, at the highest point, the given and unknown quantities are related by Equation \eqref{wROSXN}:

\begin{equation*}
    v_y^2 = v_{0y}^2 - 2 g \left(y-y_0\right)
\end{equation*}

Substituting known values leads to

\begin{equation*}
    0^2 = \left(67.6\right)^2 - 2 (9.8) \left(y - 0\right)
\end{equation*}

or, more simply, to

\begin{equation*}
    0 = 67.6^2 - 19.6 y
\end{equation*}

We can solve this equation for height $y$ as follows:

\begin{align*}
    \textbf{Add $19.6 y$} \qquad & 0 \redplus \textcolor{red}{19.6 y} = 67.6^2 - \cancel{19.6 y} \redplus \textcolor{red}{\cancel{19.6 y}}\\[1ex]
    \textbf{Simplify} \qquad & 19.6 y = 67.6^2\\[1ex]
    \textbf{Divide by 19.6} \qquad & \frac{\cancel{19.6} y}{\textcolor{red}{\cancel{19.6}}} = \frac{67.6^2}{\textcolor{red}{19.6}}\\[1ex]
    \textbf{Simplify} \qquad & y = 233
\end{align*}

Therefore, the maximum height of the firework is 233 meters.

\solutionEnd

\begin{example} \label{t57osh}
    How much time passed between the launch of the shell in Example \ref{4Dn9PW} and the explosion?
\end{example}

\Solution
Upon solving Example \ref{4Dn9PW}, we learned that initial and final vertical velocities associated with a launch to maximum height are

\begin{equation*}
    v_{0y} = \SI{67.6}{m/s} \quad \text{and} \quad v_y = \SI{0}{m/s}
\end{equation*}

Furthermore, gravitational acceleration is always $g = \SI{9.8}{m/s^2}$. Since we are now looking for how long it took the firework to reach maximum height, the unknown is time: $t =\ ?$

\vspace{1em}

The given and unknown quantities are now related by Equation \eqref{SjYaoE} as 

\begin{equation*}
    v_y = v_{0y} - gt
\end{equation*}

Substituting known values leads to

\begin{equation*}
    0 = 67.6 - 9.8 t
\end{equation*}

Let's solve this equation as follows:

\begin{align*}
    \textbf{Add $9.8 t$} \qquad & 0 \redplus \textcolor{red}{9.8 t} = 67.6 - \cancel{9.8 t} \redplus \cancel{\textcolor{red}{9.8 t}}\\[1ex]
    \textbf{Simplify} \qquad & 9.8 t = 67.6\\[1ex]
    \textbf{Divide by 9.8} \qquad & \frac{\cancel{9.8} t}{\cancel{\textcolor{red}{9.8}}} = \frac{67.6}{\textcolor{red}{9.8}}\\[1ex]
    \textbf{Simplify} \qquad & t = 6.90
\end{align*}

Therefore, from the launch of the firework to its explosion at maximum height, a total of 6.90 seconds elapsed.

\solutionEnd

\begin{example} 
    What is the horizontal displacement of the shell in Example \ref{4Dn9PW} when it explodes?
\end{example}

\Solution In this example, since we are asked to find horizontal displacement, we analyze only the horizontal motion. Assuming the firework is launched from the origin, its initial position is $x_0 = 0$. Since the initial speed and launch angle are $v_0 = \SI{70}{m/s}$ and $\theta_0 = \ang{75}$,  the horizontal component of initial velocity is

\begin{equation*}
    v_{0x} = v_0 \cos{\theta_0} = 70 \cos{(\ang{75})} = \SI{18.1}{m/s}
\end{equation*}

In horizontal motion, by Equation \eqref{dCvDf4}, velocity is a constant, meaning that the projectile, throughout its trajectory, will maintain a constant horizontal speed of 

\begin{equation*}
    v_x = v_{0x} = \SI{18.1}{m/s}
\end{equation*}


Furthermore, in Example \ref{t57osh}, we concluded that the time from launch to maximum height is $t = \SI{6.90}{s}$. Finally, the unknown we are looking for is displacement from the origin: $x =\ ?$.

\vspace{1em}

The known and unknown quantities are related by Equation \eqref{7fMAPG} as

\begin{equation*}
    x = x_0 + v_x t
\end{equation*}

Substituting known values leads to 

\begin{equation*}
    x = 0 + (18.1)(6.90) = 125 
\end{equation*}

Therefore, the firework was horizontally displaced by 125 meters.

\solutionEnd


The fact that vertical and horizontal motions are independent of each other lets us predict the range of a projectile. The \gls{range} is the horizontal distance $R$ traveled by a projectile on level ground, as illustrated in Figure \ref{6hKQ2Y}. Throughout history, people have been interested in finding the range of projectiles for practical purposes, such as aiming cannons.

\begin{center}
    \begin{tikzpicture}[
        declare function={f(\x,\yi,\vi,\thetai)=\yi + tan(\thetai)*\x - \grav*\x^2/(2*(\vi*cos(\thetai))^2);}, %equation of path
        declare function ={R(\vi,\thetai)= \vi^2*sin(2*\thetai)/\grav;}, %range
        declare function={h(\vi,\thetai)=(\vi*sin(\thetai))^2/(2*\grav);}, %maximum height
    ]
    \tikzmath{
        \grav = 9.8;
        \sf = 0.5; %scale factor for vector components
    }
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=12cm,height=5cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax={R(50,45)*1.1},
        ymin=0, ymax={h(50,45)*1.1},
        ]
        \draw[Green,thick,->] (0,0) -- ++({50*cos(45)},{50*sin(45)}) node[above left=-5pt] {\SI{50}{m/s}};;
        \draw[Green,thick,domain=0:{R(50,45)},variable=\x,samples=250] plot ({\x},{f(\x,0,50,45)}) node[black,below] {$R = \SI{225}{m}$};
        \draw[violet,thick,->] (0,0) -- ++({40*cos(45)},{40*sin(45)}) node[pos=0.93,right=2.7mm] {\SI{40}{m/s}};;
        \draw[violet,thick,domain=0:{R(40,45)},variable=\x,samples=250] plot ({\x},{f(\x,0,40,45)}) node[black,below] {$R = \SI{163}{m}$};
        \draw[red,thick,->] (0,0) -- ++({30*cos(45)},{30*sin(45)}) node[pos=0.4,right=3pt] {\SI{30}{m/s}, \ang{45}};
        \draw[red,thick,domain=0:{R(30,45)},variable=\x,samples=250] plot ({\x},{f(\x,0,30,45)})
    node[black,below] {$R = \SI{91.8}{m}$};;
        \end{axis}
    \end{tikzpicture}

    \vspace{1em}

    \begin{tikzpicture}[
        declare function={f(\x,\yi,\vi,\thetai)=\yi + tan(\thetai)*\x - \grav*\x^2/(2*(\vi*cos(\thetai))^2);}, %equation of path
        declare function ={R(\vi,\thetai)= \vi^2*sin(2*\thetai)/\grav;}, %range
        declare function={h(\vi,\thetai)=(\vi*sin(\thetai))^2/(2*\grav);}, %maximum height
    ]
    \tikzmath{
        \grav = 9.8;
        \sf = 0.5; %scale factor for vector components
    }
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=12cm,height=5cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax={R(50,45)*1.1},
        ymin=0, ymax={h(50,75)*1.1},
        ]
        \draw[Green,thick,->] (0,0) -- ++({50*cos(75)},{50*sin(75)}) node[pos=1.7,rotate=58] {\SI{50}{m/s}, \ang{75}};;
        \draw[Green,thick,domain=0:{R(50,75)},variable=\x,samples=250] plot ({\x},{f(\x,0,50,75)}) node[black,below] {$R = \SI{128}{m}$};
        \draw[violet,thick,->] (0,0) -- ++({50*cos(45)},{50*sin(45)}) node[pos=1.6,rotate=20,below=-6pt] {\SI{50}{m/s}, \ang{45}};;
        \draw[violet,thick,domain=0:{R(50,45)},variable=\x,samples=250] plot ({\x},{f(\x,0,50,45)}) node[black,below] {$R = \SI{225}{m}$};
        \draw[red,thick,->] (0,0) -- ++({50*cos(15)},{50*sin(15)}) node[pos=1.5] {\SI{50}{m/s}, \ang{15}};;
        \draw[red,thick,domain=0:{R(50,15)},variable=\x,samples=250] plot ({\x},{f(\x,0,50,15)});
        \end{axis}
    \end{tikzpicture}
    \captionsetup{type=figure,margin=1in,font=scriptsize}
    \captionof{figure}{Trajectories of projectiles on level ground. \textit{Top plot}: The greater the initial speed,$v_0$, the greater the range for a given initial angle. \textit{Bottom plot}: The effect of initial angle, $\theta_0$, on the range of a projectile with a given initial speed. Note that any combination of trajectories that add to 90 degrees will have the same range in the absence of air resistance, although the maximum heights of those paths are different.}
    \label{6hKQ2Y}
\end{center}

How does the initial velocity of a projectile affect its range? Obviously, the greater the initial speed $v_0$, the greater the range, as shown in the figure above. The initial angle $\theta_0$ also has a dramatic effect on the range. When air resistance is negligible, the range $R$ of a projectile on level ground is

\begin{equation} \label{GqvaN1}
    R = \frac{v_0^2 \sin{(2\theta_0)}}{g}
\end{equation}

where $v_0$ is the initial speed and $\theta_0$ is the initial angle relative to the horizontal. It is important to note that the range doesn't apply to problems where the initial and final $y$ position are different, or to cases where the object is launched perfectly horizontally.

\begin{example}
    \href{https://youtu.be/9Ijev8RHF04}{Brynden the Blackfish} uses a bow to launch an arrow at a speed of \SI{39}{m/s}. The arrow's trajectory looks like those shown in Figure \ref{6hKQ2Y}. If the launch angle is \ang{56}, what is the arrow's range? Ignore air resistance (which, actually, \textit{could} affect the motion of the arrow.) 
\end{example}

\Solution We are given the initial speed and launch anlge: $v_0 = \SI{39}{m/s}$ and $\theta_0 = \ang{56}$. Acceleration due to gravity always has a magnitude of $g = \SI{9.8}{m/s^2}$. 

\vspace{1em}

The unknown we are looking for is range, which is given by Equation \eqref{GqvaN1} as

\begin{equation*}
    R = \frac{v_0^2 \sin{2\theta_0}}{g}
\end{equation*}

Substituting known values leads to

\begin{equation*}
    R = \frac{39^2 \sin{(2 \cdot \ang{56})}}{9.8} = 144
\end{equation*}

Therefore, the range of the arrow is 144 meters. That's how far it can travel horizontally.

\solutionEnd

\begin{example} \label{9XXF9X}
    George drops a rock from rest from the top of an \SI{82}{m} tall cliff. Calculate how long it will take the rock to impact the ground. 
\end{example}

\Solution We are given the rock's initial velocity (``from rest''): $v_{0y} = 0$. Taking the rock's initial vertical position to be the origin, then we are also given its initial and final positions: $y_0 = 0$ and $y = \SI{-82}{m}$. Furthermore, we know the magnitude of gravitational acceleration is $g = \SI{9.8}{m/s^2}$. 

\vspace{1em}

The unknown we are looking for is time: $t =\ ?$

\vspace{1em}

The given and unknown quantities are related by Equation \eqref{36YuvF}:

\begin{equation*}
    y = y_0 + v_{0y}t - \frac{1}{2}  g t^2
\end{equation*}

Substituting given values leads to 

\begin{equation*}
    -82 = 0 + (0) t - \frac{1}{2} g t^2
\end{equation*}

or, more simply, to

\begin{equation*}
    -82 = - 4.9 t^2
\end{equation*}

Since we are solving this equation for time, $t$, we should write the unknown variable on the left:

\begin{equation*}
    -4.9 t^2 = -82
\end{equation*}

This equation is solved as follows:

\begin{align*}
    \textbf{Divide by $-4.9$} \qquad & \frac{\cancel{-4.9} t^2}{\cancel{\redminus \textcolor{red}{4.9}}} = \frac{-82}{\redminus \textcolor{red}{4.9}}\\[1ex]
    \textbf{Simplify} \qquad & t^2 = \frac{82}{4.9}\\[1ex]
    \textbf{Take square root} \qquad & \sqrt{t^2} = \sqrt{\frac{82}{4.9}}\\[1ex]
    \textbf{Simplify} \qquad & t = 4.09
\end{align*}

Therefore, dropping from rest, it takes this rock 4.09 seconds to strike the ground.

\solutionEnd


\begin{example} \label{lB7EDG}
    A cannon is aimed at a right angle to the horizontal, perfectly upwards, next to a 76-meter-tall redwood tree. The cannon fires a cannonball at a speed of \SI{42}{m/s}. How much time after launch does it take the cannonball to reach the top of the tree, if the projectile has passed maximum height and is on its way down?
\end{example}

\Solution Since the cannonball's motion is strictly vertical, we will work with equations governing vertical motion only. We are given the cannonball's initial speed: $v_{0y} = \SI{42.0}{m/s}$. Gravitational acceleration always has a magnitude of $g = \SI{9.80}{m/s^2}$. If we assume the cannonball starts at the origin, then we know its initial and final positions: $y_0 = 0$ and $y = \SI{76.0}{m}$. The unknown we are looking for is the time: $t =\ ?$.

\vspace{1em}

The given and unknown quantities are related by Equation \eqref{36YuvF}:

\begin{equation*}
    y = y_0 + v_{0y}t - \frac{1}{2}  g t^2
\end{equation*}

Substituting given values leads to 

\begin{equation*}
    76 =  0 + 42 t -\frac{1}{2}(9.8) t^2\ ,
\end{equation*}

or, more simply, to

\begin{equation*}
    76 = 42 t - 4.9 t^2
\end{equation*}

We may solve for time, $t$, as follows:

\begin{align*}
    \textbf{Add $4.9 t^2$} \qquad & 76 \redplus \textcolor{red}{4.9 t^2} = 42 t - 4.9 t^2 \redplus \textcolor{red}{4.9 t^2}\\[1ex]
    \textbf{Simplify} \qquad & 76 + 4.9 t^2 = 42 t\\[1ex]
    \textbf{Subtract $42 t$} \qquad & 76 + 4.9 t^2 \redminus \textcolor{red}{42 t} = 42 t \redminus \textcolor{red}{42 t}\\[1ex]
    \textbf{Simplify} \qquad & 76 + 4.9 t^2 - 42 t = 0
\end{align*}

The last equation may be rearranged as 

\begin{equation} \label{LScbXD}
    4.9 t^2 - 42 t + 76 = 0
\end{equation}

so that it resembles the quadratic form

\begin{equation} \label{WsGUny}
    ax^2 + bx + c = 0
\end{equation}

whose two solutions are given by the quadratic equation

\begin{equation} \label{BMWMHy}
    x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
\end{equation}

Setting the coefficients in Equation \eqref{LScbXD} to match the symbols in Equation \eqref{WsGUny} leads to 

\begin{equation*}
    a = 4.9\ , \quad b = -42\ ,\quad c = 76
\end{equation*}

Substituting these coefficients into Equation \eqref{BMWMHy} leads to the first solution of time:

\begin{equation*}
    t_1 = \frac{-b + \sqrt{b^2 - 4ac}}{2a} = \frac{-(-42) + \sqrt{(-42)^2 - 4 (4.9)(76)}}{2(4.9)} = \SI{5.98}{s}
\end{equation*}

Likewise, the second solution of time is given by

\begin{equation*}
    t_2 = \frac{-b - \sqrt{b^2 - 4ac}}{2a} = \SI{2.60}{s}
\end{equation*}

Since the second solution, $t_2 = \SI{2.60}{s}$, is less than the first, it must correspond to the cannonball reaching the top of the tree ($y=\SI{76}{m}$) on the way \textit{up}. The first solution ($t_1 = \SI{5.98}{s}$) corresponds to the projectile reaching the same point on its way back down. Therefore, the cannonball will reach the top of the tree, on its way back down, 5.98 seconds after the initial launch.

\solutionEnd


\begin{example} \label{pyn6Go}
    George, from Example \ref{9XXF9X}, this time throws a rock downwards at a speed of \SI{17.00}{m/s} from the top of the same \SI{82}{m} tall cliff. Calculate how long it will take the rock to impact the ground now.
\end{example}

\Solution We are given initial velocity: $v_{0y} = \SI{-17.00}{m/s}$. Assuming that the rock's initial vertical position is zero, then we are given the following initial and final positions: $y_0 = 0$ and $y = \SI{-82}{m}$. Also, we know the magnitude of gravitational acceleration is always $g = \SI{9.8}{m/s^2}$. 

\vspace{1em}

The unknown we are looking for is time: $t =\ ?$

\vspace{1em}

The given and unknown quantities are related by Equation \eqref{36YuvF}:

\begin{equation*}
    y = y_0 + v_{0y}t - \frac{1}{2}  g t^2
\end{equation*}

This is the same equation and unknown that was referred to in Example \ref{lB7EDG}, so the current solution will develop almost exactly like the one from Example \ref{lB7EDG}, even though the two examples involve different trajectories.


\vspace{1em}

Substituting given values leads to 

\begin{equation*}
    -82 =  0 - 17 t -\frac{1}{2}(9.8) t^2\ ,
\end{equation*}

or, more simply, to

\begin{equation*}
     -82 = - 17 t -4.9 t^2
\end{equation*}

Let's rearrange this equation as follows:

\begin{align*}
    \textbf{Add $17t$} \qquad & - 82 \redplus \textcolor{red}{17t} = -17t \redplus \textcolor{red}{17t} - 4.9 t^2 \\[1ex]
    \textbf{Simplify} \qquad & -82 + 17t = -4.9t^2\\[1ex]
    \textbf{Add $4.9 t^2$} \qquad & -82 + 17t \redplus \textcolor{red}{4.9 t^2} = -4.9t^2 \redplus \textcolor{red}{4.9 t^2}\\[1ex]
    \textbf{Simplify} \qquad & -82 + 17t + 4.9t^2 = 0
\end{align*}

The last equation is rearranged as 

\begin{equation} \label{dw27jh}
    4.9 t^2 + 17t - 82 = 0
\end{equation}

so that it fits the quadratic form $ax^2 + bx + c = 0$, which has two solutions given by the quadratic equation \eqref{BMWMHy}. Let the coefficients in Equation \eqref{dw27jh} be

\begin{equation*}
    a = 4.9\ , \quad b = 17\ ,\quad c = -82
\end{equation*}

so that the first root of time is

\begin{equation*}
    t_1 = \frac{-b + \sqrt{b^2 - 4ac}}{2a} = \frac{-17 + \sqrt{17^2 - 4(4.9)(-82)}}{2(4.9)} = \SI{2.71}{s}
\end{equation*}

The second solution is found similarly as

\begin{equation*}
    t_2 = \frac{-b - \sqrt{b^2 - 4ac}}{2a} = \SI{-6.18}{s}
\end{equation*}

Since the rock must strike the ground \textit{after} it is thrown, we take the first solution of time as the answer to this exercise: the rock strikes the ground 2.71 seconds after it is launched. 

\solutionEnd

\begin{example}
    Suppose a large rock is ejected from a volcano, as illustrated in Figure ?.??, with a speed of  \SI{25.0}{m/s} and at an angle \ang{35} above the horizontal. The rock strikes the side of the volcano at an altitude 20.0 m lower than its starting point. Calculate the time it takes the rock to follow this path.
\end{example}

\begin{center}
    \begin{tikzpicture}[
        declare function={f(\x)=\yi + tan(\thetai)*\x - \grav*\x^2/(2*(\vi*cos(\thetai))^2);},
        declare function={t(\x)=\x/(\vi*cos(\thetai));}
    ]
    \tikzmath{
        \grav = 9.8;
        \vi = 25;
        \yi = 0.0;
        \thetai = 35.0;
        \viy = \vi*sin(\thetai);
        \ymax = (\viy^2)/(2*\grav);
        \tmax = 2*\ymax/\viy;
        \vix = \vi*cos(\thetai);
        \xmax = \vix*3.96; %t = 3.96 s
        \sf = 0.5; %scale factor for vector components
    }
    \pgfplotsset{compat=1.11}
        \begin{axis}[width=12cm,height=7cm,ticks=none,
        axis lines = center,
        clip=false,
        ylabel = $y$,
        xlabel = $x$,
        xmin=0, xmax=100,
        ymin=-30, ymax=10,
        ]
        \draw[domain=0:\xmax,variable=\x,samples=250] plot ({\x},{f(\x)});
        \draw[ultra thick,violet,->] (0,0) -- ++(axis direction cs: {\sf*\vix},{\sf*\viy}) node[above=-1pt] {$\SI{25}{m/s}$};
        \draw (5,0) arc (0:35:5) node[pos=0.1,right=-1pt] {\ang{35}};
        \draw[dashed] (\xmax,-20) -- ++(axis direction cs: 0,20) node[pos=0.5,right] {-\SI{20}{m}};
        \draw[fill=black] (\xmax,-20) circle (2pt);
        \draw (\xmax,-20) -- ++(axis direction cs: 12,0);
        \draw[,ultra thick,->,violet] (\xmax,-20) -- ++(axis direction cs: {\sf*\vix},{\sf*(\viy -\grav*3.96)}) node[below=-2pt] {$v$};
        \draw (\xmax+5,-20) arc (0:-48:5) node[pos=0.1,right=-1pt] {$\theta$};
        \end{axis}
    \end{tikzpicture}
    \captionsetup{type=figure,margin=1in,font=scriptsize}
    \captionof{figure}{The diagram represents the projectile motion of a large rock from a volcano.}
\end{center}

\Solution Recall the most important concept in projectile motion: horizontal and vertical motions are independent. Since they don't influence one another, we can analyze them separately. For this problem, we can ignore calculations related to the horizontal motion and focus only on the vertical motion. As such, the solution to this example will develop like Examples \ref{lB7EDG} and \ref{pyn6Go}. 

\vspace{1em}

We are given the initial speed and launch angle: $v_0 = \SI{25.0}{m/s}$ and $\theta_0 = \ang{35}$. If we set the origin to be the rock's initial position, then the initial and final vertical positions are $y_0 = 0$ and $y = \SI{-20.0}{m}$. The magnitude of gravitational acceleration is $g = \SI{-9.8}{m/s^2}$.

\vspace{1em}

Since we are analyzing vertical motion only, we get the $y$-component of initial velocity by

\begin{equation*}
    v_{0y} = v_0 \sin{\theta_0} = 25 \sin{(\ang{35})} = 14.34
\end{equation*}

 The unknown quantity we are looking for is time: $t =\ ?$ These given and unknown quantities are related by Equation \eqref{36YuvF}

\begin{equation*}
    y = y_0 + v_{0y}t - \frac{1}{2} g t^2
\end{equation*}

At this point, this example is solved like the solutions to Examples \ref{lB7EDG} and \ref{pyn6Go}. As shown previously, we can define the coefficients $a$, $b$, and $c$ to fit the given quantities from this example. Then, applying the quadratic formula \eqref{BMWMHy} with those coefficients leads to the two solutions (roots) of time:

\begin{equation*}
    t_1 = \SI{3.96}{s} \quad \text{and} \quad t_2 = \SI{-1.03}{s}
\end{equation*}

Choosing the positive solution, we conclude that the rock will strike the ground 3.96 seconds after launch.


\subsection{Uniform Circular Motion} \label{uPOucv}

Circular motion is when an object moves in a circular path. Examples of circular motion include a race car speeding around a circular curve, a toy attached to a string swinging in a circle around your head, or the circular loop-the-loop on a roller coaster. The simplest case of circular motion is \gls{uniform circular motion}, where an object travels a circular path at a constant speed. Note that, unlike speed, the linear velocity of an object in circular motion is constantly changing because it is always changing direction. We know that acceleration is a change in velocity, either in magnitude or in direction or both. Therefore, an object undergoing uniform circular motion is always accelerating, even though the magnitude of its velocity is constant.

\vspace{1em}

You experience this acceleration yourself every time you ride in a car while it turns a corner. If you hold the steering wheel steady during the turn and move at a constant speed, you are executing uniform circular motion. What you notice is a feeling of sliding (or being flung, depending on the speed) away from the center of the turn. This isn't an actual force that is acting on you---it only happens because your body wants to continue moving in a straight line (as per Newton's first law) whereas the car is turning off this straight-line path. Inside the car it appears as if you are forced away from the center of the turn. This fictitious force is known as the \gls{centrifugal force}. The sharper the curve and the greater your speed, the more noticeable this effect becomes.

\begin{center}
\begin{tikzpicture}
    \begin{axis}[
        width=7cm,height=7cm,
        axis line style={draw=none},
        ticks=none,
        axis lines=middle,
        xmin=-1,xmax=1,
        ymin=-1,ymax=1,
        clip=false,
    ]
    \draw[gray] (0,0) circle (1);
    \fill (0,0) circle (1pt);
    \fill ({cos(45)},{sin(45)}) circle (3pt) node[above right] {$m$};
    \begin{scope}[shift={(axis direction cs: 0.04,-0.04)}]
        \draw[dashed,<->] (0,0) -- ({cos(45)},{sin(45)}) node[pos=0.5,below right=-1pt] {$r$};
    \end{scope}
    \draw[Green,very thick,->] ({cos(45)},{sin(45)}) -- ++(axis direction cs: -0.3,-0.3) node[black,above=5pt,pos=1.2] {$a_{\text{c}}$};
    \draw[red,very thick,->] ({cos(45)},{sin(45)}) -- ++(axis direction cs: -0.6,0.6) node[black,pos=1.1] {$v$};
\end{axis}
\end{tikzpicture}
\captionsetup{type=figure,margin=1in,font=scriptsize}
\captionof{figure}{Uniform circular motion, the radius of curvature and vectors for tangential velocity and centripetal acceleration. Compare this with the Ladybug Motion Simulation in Exercise \ref{N066SF}.}
\label{4QmpMJ}
\end{center}

Other examples of (approximate) uniform circular motion include the \href{https://youtu.be/3NYlYN-QTig}{hammer throw}, a bicyclist rounding a turn on a curved path, and the Earth's orbit around the Sun. The \gls{tangential velocity} ($\vec{v}$) is the instantaneous linear velocity of the object in circular motion. If the circular motion is abruptly cut off (as in the hammer throw sport), the object moves in a straight line in the direction of the tangential velocity. \Gls{tangential speed} is the magnitude of the tangential velocity, with no regard for the direction. The \gls{radius of curvature} ($r$) is the distance between the center of the circle and the path. We call the acceleration of an object moving in uniform circular motion the \gls{centripetal acceleration} ($a_c$) because centripetal means \textit{center seeking}.

\vspace{1em}

Now that we know that the direction of centripetal acceleration is toward the center of rotation, let's discuss the magnitude of centripetal acceleration. For an object traveling at speed $v$ in a circular path with radius $r$, the magnitude of centripetal acceleration is


\begin{equation} \label{5vsZ8y}
    a_{\text{c}} = \frac{v^2}{r}\ 
\end{equation}

\begin{center}
    \begin{tabular}{cl|cl}
    \hline
    \textbf{Symbol} & \textbf{Quantity} & \textbf{Unit} & \textbf{Unit Symbol}  \\
    \hline\hline
    \rule{0pt}{2.5ex}
        $a_{\text{c}}$ & centripetal acceleration & meter per second squared & \SI{}{m/s^2}\\
        $v$ & tangential speed & meter per second & \SI{}{m/s} \\
        $r$ & radius & meter & \SI{}{m}\\
    \hline
    \end{tabular}
\end{center}

Centripetal acceleration is greater at high speeds and in sharp curves (smaller radius), as you may have noticed when driving a car, because the car actually pushes you toward the center of the turn. But it is a bit surprising that ac is proportional to the speed squared. This means, for example, that the acceleration is four times greater when you take a curve at \SI{20}{m/s} than at \SI{10}{m/s}.

\vspace{1em}

Because an object in uniform circular motion undergoes acceleration (by changing the direction of motion but not the speed), we know from Newton's second law of motion that there must be a net external force acting on the object. Since the magnitude of the acceleration is constant, so is the magnitude of the net force, and since the acceleration points toward the center of the rotation, so does the net force.

\vspace{1em}

Any force or combination of forces can cause a centripetal acceleration. Just a few examples are the tension in the rope on a tether ball, the force of Earth's gravity on the Moon, the friction between a road and the tires of a car as it goes around a curve, or the normal force of a roller coaster track on the cart during a loop-the-loop.

\vspace{1em}

The component of any net force that causes circular motion is called a \gls{centripetal force}. When the net force is equal to the centripetal force, and its magnitude is constant, uniform circular motion results. The direction of a centripetal force is toward the center of rotation, the same as for centripetal acceleration. According to Newton's second law of motion, a net force causes the acceleration of mass according to 

\begin{equation*}
    F_{\text{net}} = ma
\end{equation*}

For uniform circular motion, the acceleration is centripetal acceleration: 

\begin{equation*}
    a = a_c
\end{equation*}

Therefore, the magnitude of centripetal force, $F_c$, is $F_c = m a_c$. By using Equation \eqref{5vsZ8y} for the magnitude of centripetal acceleration, $a_c = v^2/r$, we get an expression involving the magnitude of the centripetal force in terms of tangential speed:

\begin{equation} \label{4H9EJq}
    F_c = m a_c = \frac{m v^2}{r}
\end{equation}

\begin{center}
    \begin{tabular}{cl|cl}
    \hline
    \textbf{Symbol} & \textbf{Quantity} & \textbf{Unit} & \textbf{Unit Symbol}  \\
    \hline\hline
    \rule{0pt}{2.5ex}
        $F_{\text{c}}$ & centripetal force & newton & N\\
        $m$ & mass & kilogram & \SI{}{kg}\\
        $v$ & tangential speed & meter per second & \SI{}{m/s} \\
        $r$ & radius & meter & \SI{}{m}\\
    \hline
    \end{tabular}
\end{center}

Newton's second law also states that the object will accelerate in the same direction as the net force. By definition, the centripetal force is directed towards the center of rotation, so the object will also accelerate towards the center. A straight line drawn from the circular path to the center of the circle will always be perpendicular to the tangential velocity. 

\begin{example}
    A car follows a curve of radius \SI{500}{m} at a speed of \SI{25.0}{m/s} (about \SI{56}{mph}). What is the magnitude of the car's centripetal acceleration?
\end{example}

\Solution We are given the radius of curvature and the tangential velocity:

\begin{equation*}
    r = \SI{500}{m} \quad \text{and} \quad v = \SI{25}{m/s}
\end{equation*}

The unknown is centripetal acceleration: $a_c =\ ?$. By Equation \eqref{5vsZ8y},

\begin{equation*}
    a_c = \frac{v^2}{r} = \frac{25^2}{500} = 1.25 
\end{equation*}

The centripetal acceleration on the car is \SI{1.25}{m/s} (about 13\% of $g$, the acceleration due to gravity).

\solutionEnd

\begin{example}
    What is the centripetal force exerted on a \SI{1600}{kg} car that rounds a \SI{100}{m} radius curve at \SI{12}{m/s}?
\end{example}

\Solution We are given mass, radius, and speed:

\begin{equation*}
    m = \SI{1600}{kg}\ , \quad 
    r = \SI{100}{m}\ , \quad
    v = \SI{12}{m/s}
\end{equation*}

The unknown is centripetal force: $F_c =\ ?$ The given and unknown values are related Equation \eqref{4H9EJq}:

\begin{equation*}
    F_c = \frac{m v^2}{r} = \frac{(1600)(12^2)}{100} = 2304
\end{equation*}

The centripetal force is 2304 newtons inward. 

\solutionEnd

\subsection{Exercises}

\subsubsection*{\nameref{tMF7JY}}

\begin{exercise} \label{90PABl}
    What is the angle between the $x$ and $y$ components of a vector?
\end{exercise}

\begin{exercise}
    Horizontal and vertical motions of a projectile are independent of each other. What does this mean?
\end{exercise}

\begin{exercise} \label{YYBBp5}
    Using the conventional choice for positive and negative axes described in the text, what is the $y$-component of the acceleration of an object experiencing projectile motion?
\end{exercise}

\begin{exercise}  \label{agp59Z}
    Two identical items, object 1 and object 2, are dropped from the top of a \SI{50.0}{m} building. Object 1 is dropped with an initial velocity of \SI{0}{m/s}, while object 2 is thrown straight downward with an initial velocity of \SI{13.0}{m/s}. What is the difference in time, in seconds rounded to the nearest tenth, between when the two objects hit the ground?
\end{exercise}

\begin{exercise} \label{Xu3FK9}
    A water balloon cannon is fired at \SI{30}{m/s} at an angle of \ang{50} above the horizontal. How far away will it fall?
\end{exercise}

\begin{exercise} \label{piJ9kJ}
    A person wants to fire a water balloon cannon such that it hits a target \SI{100}{m} away. If the cannon can only be launched at \ang{45} above the horizontal, what should be the initial speed at which it is launched?
\end{exercise}

\begin{exercise} \label{sy5RGT}
    After a projectile is launched in the air, in which direction does it experience constant, non-zero acceleration, ignoring air resistance?
\end{exercise}

\begin{exercise} \label{N0Bwou}
    A ball is thrown in the air at an angle of \ang{40}. If the maximum height it reaches is \SI{10}{m}, what must be its initial speed?
\end{exercise}

\begin{exercise} \label{lhjg95}
    A large rock is ejected from a volcano with a speed of \SI{30}{m/s} and at an angle \ang{60} above the horizontal. The rock strikes the side of the volcano at an altitude of \SI{10.0}{m} lower than its starting point. Calculate the horizontal displacement of the rock.
\end{exercise}

\begin{exercise} \label{wlbvvu}
    How can you express the velocity, $v$, of a projectile in terms of its initial velocity, $v_0$, acceleration, $g$, and time, $t$?
\end{exercise}

% \begin{exercise}
%     In the equation for the maximum height of a projectile, what does $v_{0y}$ stand for?

% \begin{equation*}
%     h = \frac{v_{0y}^2}{2g}
% \end{equation*}
% \end{exercise}

\begin{exercise} \label{xj6gln}
    True or False? Range is defined as the maximum vertical distance travelled by a projectile. 
\end{exercise}

\begin{exercise} \label{5wfdN0}
    For what angle of a projectile is its range equal to zero?
\end{exercise}

\begin{exercise} \label{DnqDpu}
    Ignoring drag, what is the $x$-component of the acceleration of a projectile? Why?
\end{exercise}

\begin{exercise} \label{QsBUvP}
    What is the optimum angle at which a projectile should be launched in order to cover the maximum distance?
\end{exercise}

\subsubsection*{Uniform Circular Motion}

\begin{exercise} \label{N066SF}
    Bring Figure \ref{4QmpMJ} to life by taking the following steps:

    \begin{enumerate}
    \setlength\itemsep{0.1ex}
        \item Access the OpenStax Simulations (\href{https://veillette.github.io/simulations/}{click here}).
        \item Click the simulation called ``Ladybug Motion.''
        \item Under \texttt[Green]{Vectors}, check the boxes for \texttt[Green]{Show Velocity} and \texttt[Green]{Show Acceleration}. 
        \item Towards the bottom, click \texttt[red]{Velocity}. Then, under \texttt[Green]{Remote Control}, drag the red button until you generate a red arrow about 1 centimeter in length. The Ladybug will start moving.
        \item When the Ladybug is in motion, locate the \texttt[Green]{Motion} options, and select \texttt[Green]{Circular}. The Ladybug is now undergoing uniform circular motion.
        \item Draw a sketch of the Ladybug and its velocity and acceleration vectors. Compare your sketch to Figure \ref{4QmpMJ}. 
    \end{enumerate}
\end{exercise}

\begin{exercise} \label{x2YCzr}
    What is meant by the word \textit{centripetal}?
\end{exercise}

\begin{exercise} \label{S4rGIt}
    (a) What is the direction of the velocity of an object in uniform circular motion? (b) What is the direction of the acceleration?
\end{exercise}

\begin{exercise} \label{aVu3ZA}
    Suppose you have an object tied to a rope and are rotating it over your head in uniform circular motion. If you increase the length of the rope, would you have to apply more or less force to maintain the same speed?
\end{exercise}

\begin{exercise} \label{kTvEZX}
    What is the centripetal acceleration felt by the passengers of a car moving at \SI{12}{m/s} along a curve with radius \SI{2.0}{m}?
\end{exercise}


\begin{exercise} \label{MJm5VE}
    Find the frictional force between the tires and the road that allows a \SI{1000}{kg} car traveling at \SI{30}{m/s} to round a \SI{20}{m} radius curve.
\end{exercise}

\begin{exercise} \label{fsjtBV}
    Which of the following quantities is constant in uniform circular motion? Speed, velocity, acceleration, displacement.
\end{exercise}

\begin{exercise} \label{GD8Yff}
    List the three quantities that influence centripetal force.
\end{exercise}

\begin{exercise} \label{aEtlVz}
    An increase in the magnitude of the \rule{2cm}{0.15mm} causes a \textit{reduction} in the centripetal force.
\end{exercise}

\begin{exercise} \label{5lauZ6}
    What happens to centripetal acceleration as the radius of curvature \textit{decreases} and the speed is held constant? Why?
\end{exercise}

\begin{exercise} \label{KB7t4k}
    Why do we experience more sideways acceleration while driving around sharper curves?
\end{exercise}

\begin{exercise} \label{jb3iGO}
    Is centripetal acceleration a vector or a scalar?
\end{exercise}

\begin{exercise} \label{TUxTQh}
    What are the units of centripetal acceleration?
\end{exercise}

\begin{exercise} \label{9YKhPo}
    What are the units of centripetal force?
\end{exercise}

\begin{exercise} \label{wS6Htw}
    What is the angle formed between the vectors of tangential velocity and centripetal force?
\end{exercise}

\begin{exercise} \label{barV9C}
    What is the angle formed between the vectors of centripetal acceleration and centripetal force?
\end{exercise}

\begin{exercise} \label{M6NkHB}
    As the mass of an object in uniform circular motion increases, what happens to the centripetal force required to keep it moving at the same speed?
\end{exercise}

\begin{exercise} \label{IAjJBC}
    Is an object in uniform circular motion accelerating? Why or why not?
\end{exercise}

\begin{exercise} \label{G8I4Up}
    An object is in uniform circular motion. Suppose the centripetal force was removed. In which direction would the object now travel?
\end{exercise}

\begin{exercise} \label{Kv8Old}
An 89-kg system consisting of a bicycle and rider are rounding a turn at \SI{16}{m/s} on a path with a radius of curvature of \SI{25}{m}. (a) What is the centripetal acceleration on the system? (b) What is the centripetal force?
\end{exercise}


\clearpage


\subsection{Answers to Select Exercises}
\ref{90PABl}. \ang{90}\\
\ref{YYBBp5}. \SI{-9.8}{m/s^2}\\
\ref{agp59Z}. \SI{1.1}{s}\\
\ref{Xu3FK9}. \SI{90.4}{m}\\
\ref{piJ9kJ}. \SI{31.3}{m/s}\\
\ref{sy5RGT}. Vertical (or $y$)\\
\ref{N0Bwou}. \SI{21.8}{m/s}\\
\ref{lhjg95}. \SI{84.9}{m}\\
\ref{wlbvvu}. See Equation \eqref{SjYaoE}\\
\ref{xj6gln}. Range is horizontal.\\
\ref{5wfdN0}. \ang{0} or \ang{90}\\
\ref{DnqDpu}. \SI{0}{m/s^2}. Gravitational acceleration is vertical only.\\
\ref{QsBUvP}. \ang{45}\\
\ref{x2YCzr}. Center-seeking\\
\ref{S4rGIt}. (a) tangential \hspace{1em} (b) radially inward.\\
\ref{aVu3ZA}. Less force. In Equation~(\ref{qzzym8}), radius (the length of the rope) is in the denominator of the fraction. This means that force and radius are \textit{inversely proportional}: if one goes up, the other goes down. By \textit{increasing} the radius (the length of the rope), the centripetal force must \textit{decrease}.
\ref{kTvEZX}. \SI{72}{m/s^2}\\
\ref{MJm5VE}. \SI{45000}{N}\\
\ref{fsjtBV}. Speed. The others are vectors whose directions may change.\\
\ref{GD8Yff}. Mass, speed, and radius. See Eq.~(\ref{qzzym8}).\\
\ref{aEtlVz}. radius of curvature\\
\ref{5lauZ6}. Acceleration is inversely proportional to radius (Eq.~\ref{5vsZ8y}). Therefore, assuming speed stays the same, as radius \textit{decreases} (as the circle gets smaller), acceleration \textit{increases}. See, also, Exercise \ref{aVu3ZA}.\\
\ref{KB7t4k}. Sharper curve means smaller radius of curvature. As radius decreases, acceleration increases. See Exercise \ref{5lauZ6}.\\
\ref{jb3iGO}. vector\\
\ref{TUxTQh}. meters per second squared (\SI{}{m/s^2})\\
\ref{9YKhPo}. newton (N)\\
\ref{wS6Htw}. \SI{90}{\degree}\\
\ref{barV9C}. \SI{0}{\degree}\\
\ref{M6NkHB}. Centripetal force must increase. Equation (\ref{qzzym8}) suggests that force and mass are \textit{directly proportional}: if one increases, the other goes increases too. \\
\ref{IAjJBC}. Yes. Although the object is not speeding up or slowing down, the direction of the velocity vector continuously changes. \\
\ref{G8I4Up}. In the direction of the tangential velocity at the instant the force was removed.\\
\ref{Kv8Old}. (a) \SI{10.24}{m/s^2} \hspace{1em} (b) \SI{911}{N}\\


\end{document}